High resolution, annual maps of the characteristics of smallholder-dominated croplands at national scales

### 2021-03-05 inst/manuscript/manuscript.Rmd
Near-final draft of MS and SI

These problems arising from cropland characteristics are increasingly being overcome due to technological advances. Recent advances in satellite technology have increased the coverage of high (<5 m) or near-high (10 m) resolution imagery with weekly to near-daily return intervals [@DruschSentinel2ESAOptical2012; @McCabefutureEarthobservation2017]. This high spatial *and* temporal resolution addresses the sensor-scale mismatch, and more effectively captures the intra-seasonal dynamics of cropland, which helps classifiers distinguish cropland from surrounding cover types [@Debatsgeneralizedcomputervision2016; @Defournyrealtimeagriculturemonitoring2019]. On top of this, advances in cloud computing [@GorelickGoogleEarthEngine2017], the opening of image archives [@WulderglobalLandsatarchive2016], and next generation machine learning approaches [@MaxwellImplementationmachinelearningclassification2018; @ZhuDeeplearningremote2017] are placing large volumes of these moderate to near-high resolution imagery together with the computational and algorithmic resources necessary to classify them at scale. These capabilities are aleady being used to create a new generation of higher resolution (10-30 m) cropland and landcover maps for Africa and other regions [@ESAESACCILAND; @LesivEvaluationESACCI2017; @XiongNominal30mCropland2017].

@@

We demonstrate this approach to map cropland in Ghana, a country where agriculture is predominantly small-scale and comprises a broad mix of climates and agricultural systems, including large areas where shifting agricultural practices predominate [@KansangaTraditionalagriculturetransition2019; @SambergSubnationaldistributionaverage2016]. These factors make Ghana emblematic of the substantial challenges facing cropland mapping over the broader region.

@@

## Assessing the characteristics of Ghanaian cropland
Using the final mapped results, we calculated the estimated area of arable cropland in Ghana, as well as the average size and total number of fields in the different AOIs. We used the map reference sample to calculate adjusted area estimates and confidence intervals for each map class, and used the differences between labellers' polygons and segmented boundaries at validation sites to calculate bias-adjusted estimates of mean field sizes and the total number of fields.

# Results
```{r, echo=FALSE, message=FALSE, warning=FALSE}
tiles <- sf::read_sf(system.file("extdata/ghana_tiles.geojson", 
                                 package = "activemapper"))
km2 <- round(as.numeric(units::set_units(sum(sf::st_area(tiles)), "km2")))
km2 <- fmt(km2)
```
We developed maps of Ghana's arable croplands, covering an area of `r km2` km$^2$ that included portions of neighboring countries overlapped by images tiles. Ghana has several agro-ecologically distinct regions, ranging from the primarily grain and vegetable crop producing regions in the northern savannas to tree crop-dominated system in the forested southwest, where cocoa and oil palm are among the dominant crops. For these latter regions, we did not attempt to classify tree crops, and instead mapped clearings that potentially contain field crops or newly felled or recently replanted tree crops. We made this decision because PlanetScope's resolution is not high enough for labellers to distinguish many tree crops from surrounding forest, and the boundaries of many tree crops (e.g. cocoa) are often not visible.

@@

### Validation performance
```{r, echo=FALSE, warning=FALSE, message=FALSE}
data("iteration_metrics")
delta_func <- function(x, rnd = 4) round((x - lag(x)) / lag(x), rnd)

@@

The potential impact of label errors on map quality was assessed in four AOIs (1, 2, 8, and 15). The results of these tests showed that the average accuracy, AUC, and F1 scores for models trained with the consensus labels were respectively `r metdiffs$Accuracy$cons`, `r metdiffs$AUC$cons`, and `r metdiffs$F1$cons` (Figure \ref{fig:trainingimpact}). Performance metrics from consensus-trained models were just `r metdiffs$AUC$chdelt` - `r metdiffs$Accuracy$chdelt` percent higher than those models trained with the most accurate individuals' labels (accuracy = `r metdiffs$Accuracy$high`'; AUC = `r metdiffs$AUC$high`; F1 = `r metdiffs$F1$high`), but were `r metdiffs$AUC$cldelt` - `r metdiffs$Accuracy$cldelt` higher than models trained with the least accurate individual labels (accuracy = `r metdiffs$Accuracy$low`; AUC = `r metdiffs$AUC$low`; F1 = `r metdiffs$F1$low`).

@@

```{r mainmap, echo = FALSE, out.width="90%", fig.cap="The distribution of arable croplands in Ghana. The main map shows the percentage of croplands in each 0.005 degree grid cell, derived from the predicted cropland probabilities. The insets  on the margins illustrate predicted probabilities (top map in each couplet) at original image resolution (0.000025 degrees) and segmented field boundaries overlaid on the dry season PlanetScope composite, for four separate tiles. Each tile's position is shown on the main map, and is color-coded to the boundary lines around its corresponding inset.", fig.align='center', message=FALSE}
knitr::include_graphics('figures/figure7.png')
```

@@

Both the per-pixel and vectorized maps, when combined with the map reference sample, enable separate estimates of the total extent of arable croplands in Ghana. The cropland extent estimated from the vectorized map is `r fmt(segas[2])` km$^2$ (with a margin of error of `r fmt(segma[1])` km$^2$), or `r segcipct[1]` (`r paste(segcipct[2:3], collapse = "-")`%) of the mapped area. The estimate based on the per pixel map is `r fmt(rfas[2])` km$^2$ (margin of error = `r fmt(rfma[1])` km$^2$), or `r rfcipct[1]` (`r paste(rfcipct[2:3], collapse = "-")`%) of area.

@@

These results demonstrate a capability for mapping the characteristics of smallholder-dominated cropping systems at high spatial resolution, annual time steps, and national scales. The resulting maps provide an updated and more granular view of the distribution and extent of arable croplands in Ghana, complementing existing national to regional land cover maps derived from moderate resolution sensors [@ESAESACCILAND; @hackmanNewLandcoverMaps2017; @XiongNominal30mCropland2017]. This prior work found that cropland covered `r round(4.62 / 23.86 * 100, 1)` [@XiongNominal30mCropland2017] to 32% [@hackmanNewLandcoverMaps2017] of Ghana in 2015, whereas our 2018 maps have cropland cover of 16.1-23.2% (Table \ref{tab:mapaccuracy}), and our map reference sample-based estimates finds `r segcipct[1]`-`r rfcipct[1]`% cover. Our results thus suggest that Ghana's arable cropland is less than previously estimated, but the difference is perhaps attributable to our use of a cropland definition that excluded longer fallows and abandoned fields, which in some regions can account for over half of the total area that could be counted as cropland [@Tongforgottenlanduse2020].

@@

Image quality was another issue, although primarily in the forested AOIs, where frequent cloud cover reduced the number of available images in all seasons, resulting in composites with more brightness artifacts and blur (Figure \ref{fig:imqual}). This impacted labellers' abilities to discern arable fields, and doubtless affected model predictions. There is little to be done to mitigate these errors, short of confining imagery to the less cloudy dry season, which may further undermine model performance, given the importance of multi-temporal imagery for cropland classification [@Debatsgeneralizedcomputervision2016; @Defournyrealtimeagriculturemonitoring2019]. Composite quality could be improved by using imagery from the same seasons over multiple years, but this would undermine the goal of developing annual maps, while the dynamism of the croplands would blur field boundaries within the imagery.

@@

# Acknowledgements
The primary support for this work was provided by Omidyar Network's Property Rights Initiative, now PlaceFund. Additional support was provided by NASA (80NSSC18K0158), and the National Science Foundation (SES-1801251; SES-1832393). Computing support was provided by the AWS Cloud Credits for Research program and the Amazon Sustainability Data Initiative. Azavea provided significant contributions in engineering the machine learning pipeline. We thank Meridia for providing initial information about local cropping systems and the characteristics of fields, and Radiant Earth Foundation for advice and guidance regarding machine learning best practices.

>>>>

These problems arising from cropland characteristics are increasingly being overcome due to technological advances. Recent advances in satellite technology have increased the coverage of high (<5 m) or near-high (10 m) resolution imagery with weekly to near-daily return intervals [@DruschSentinel2ESAOptical2012; @McCabefutureEarthobservation2017]. This high spatial *and* temporal resolution addresses the sensor-field scale mismatch, and more effectively captures the intra-seasonal dynamics of cropland, which helps classifiers distinguish cropland from surrounding cover types [@Debatsgeneralizedcomputervision2016; @Defournyrealtimeagriculturemonitoring2019]. On top of this, advances in cloud computing [@GorelickGoogleEarthEngine2017], the opening of image archives [@WulderglobalLandsatarchive2016], and next generation machine learning approaches [@MaxwellImplementationmachinelearningclassification2018; @ZhuDeeplearningremote2017] are placing large volumes of these moderate to near-high resolution imagery together with the computational and algorithmic resources necessary to classify them at scale. These capabilities are aleady being used to create a new generation of higher resolution (10-30 m) cropland and landcover maps for Africa and other regions [@ESAESACCILAND; @LesivEvaluationESACCI2017; @XiongNominal30mCropland2017].

@@

We demonstrate this approach to map cropland in Ghana, a country where agriculture is predominantly small-scale and comprises a broad mix of climates and agricultural systems, including large areas where shifting agricultural practices predominate [@KansangaTraditionalagriculturetransition2019; @SambergSubnationaldistributionaverage2016].

@@

## Assessing the characteristics of Ghanaian cropland
Using the final mapped results, we calculated the estimated area of cropland in Ghana, as well as the average size and total number of fields in the different AOIs. We used the map reference sample to calculate adjusted area estimates and confidence intervals for each map class, and used the differences between labellers' polygons and segmented boundaries at validation sites to calculate bias-adjusted estimates of mean field sizes and the total number of fields.

# Results
```{r, echo=FALSE, message=FALSE, warning=FALSE}
tiles <- sf::read_sf(system.file("extdata/ghana_tiles.geojson", 
                                 package = "activemapper"))
km2 <- round(as.numeric(units::set_units(sum(sf::st_area(tiles)), "km2")))
km2 <- fmt(km2)
```
We developed maps of Ghana's croplands, covering an area of `r km2` km$^2$ that included portions of neighboring countries overlapped by images tiles. Ghana has several agro-ecologically distinct regions, ranging from the primarily grain and vegetable crop producing regions in the northern savannas to tree crop-dominated system in the forested southwest, where cocoa and oil palm are among the dominant crops. For these latter regions, we did not attempt to classify tree crops, and instead mapped clearings that potentially contain field crops or newly felled or recently replanted tree crops. We made this decision because PlanetScope's resolution is not high enough for labellers to distinguish many tree crops from surrounding forest, and the boundaries of many tree crops (e.g. cocoa) are often not visible.

@@

### Performance gains during active learning
```{r, echo=FALSE, warning=FALSE, message=FALSE}
data("iteration_metrics")
delta_func <- function(x, rnd = 4) round((x - lag(x)) / lag(x), rnd)

@@

The potential impact of label errors on map quality was assessed in four AOIs (1, 2, 8, and 15). The results of these tests showed that the average accuracy, AUC, and F1 scores for models trained with the consensus labels were respectively `r metdiffs$Accuracy$cons`, `r metdiffs$AUC$cons`, and `r metdiffs$F1$cons` (Figure \ref{fig:trainingimpact}). Performance metrics from consensus-trained models were just `r metdiffs$AUC$chdelt` - `r metdiffs$Accuracy$chdelt` percent higher than those models trained with the most accurate individuals' labels (accuracy = `r metdiffs$Accuracy$high`; AUC = `r metdiffs$AUC$high`; F1 = `r metdiffs$F1$high`), but were `r metdiffs$AUC$cldelt` - `r metdiffs$Accuracy$cldelt` higher than models trained with the least accurate individual labels (accuracy = `r metdiffs$Accuracy$low`; AUC = `r metdiffs$AUC$low`; F1 = `r metdiffs$F1$low`).

@@

```{r mainmap, echo = FALSE, out.width="90%", fig.cap="The distribution of croplands in Ghana. The main map shows the percentage of croplands in each 0.005 degree grid cell, derived from the predicted cropland probabilities. The insets  on the margins illustrate predicted probabilities (top map in each couplet) at original image resolution (0.000025 degrees) and segmented field boundaries overlaid on the dry season PlanetScope composite, for four separate tiles. Each tile's position is shown on the main map, and is color-coded to the boundary lines around its corresponding inset.", fig.align='center', message=FALSE}
knitr::include_graphics('figures/figure7.png')
```

@@

Both the per-pixel and vectorized maps, when combined with the map reference sample, enable separate estimates of the total extent of croplands in Ghana. The cropland extent estimated from the vectorized map is `r fmt(segas[2])` km$^2$ (with a margin of error of `r fmt(segma[1])` km$^2$), or `r segcipct[1]` (`r paste(segcipct[2:3], collapse = "-")`%) of the mapped area. The estimate based on the per pixel map is `r fmt(rfas[2])` km$^2$ (margin of error = `r fmt(rfma[1])` km$^2$), or `r rfcipct[1]` (`r paste(rfcipct[2:3], collapse = "-")`%) of area.

@@

These results demonstrate a capability for mapping the characteristics of smallholder-dominated cropping systems at high spatial resolution, annual time steps, and national scales. The resulting maps provide an updated and more granular view of the distribution and extent of croplands in Ghana, complementing existing national to regional land cover maps derived from moderate resolution sensors [@ESAESACCILAND; @hackmanNewLandcoverMaps2017; @XiongNominal30mCropland2017]. This prior work found that cropland covered `r round(4.62 / 23.86 * 100, 1)` [@XiongNominal30mCropland2017] to 32% [@hackmanNewLandcoverMaps2017] of Ghana in 2015, whereas our 2018 maps have cropland cover of 16.1-23.2% (Table \ref{tab:mapaccuracy}), and our map reference sample-based estimates finds `r segcipct[1]`-`r rfcipct[1]`% cover. Our results thus suggest that Ghana's cropland is less than previously estimated, but the difference is perhaps attributable to our use of a cropland definition that excluded longer fallows and abandoned fields, which in some regions can account for over half of the total area that could be counted as cropland [@Tongforgottenlanduse2020].

@@

Image quality was another issue, although primarily in the forested AOIs, where frequent cloud cover reduced the number of available images in all seasons, resulting in composites with more brightness artifacts and blur (Figure \ref{fig:imqual}). This impacted labellers' abilities to discern fields, and doubtless affected model predictions. There is little to be done to mitigate these errors, short of confining imagery to the less cloudy dry season, which may further undermine model performance, given the importance of multi-temporal imagery for cropland classification [@Debatsgeneralizedcomputervision2016; @Defournyrealtimeagriculturemonitoring2019]. Composite quality could be improved by using imagery from the same seasons over multiple years, but this would undermine the goal of developing annual maps, while the dynamism of the croplands would blur field boundaries within the imagery.

@@

# Acknowledgements
The primary support for this work was provided by Omidyar Network's Property Rights Initiative, now PlaceFund. Additional support was provided by NASA (80NSSC18K0158), and the National Science Foundation (SES-1801251; SES-1832393). Computing support was provided by the AWS Cloud Credits for Research program and the Amazon Sustainability Data Initiative. Azavea provided significant contributions in engineering the machine learning pipeline. We thank Meridia for providing initial information about local cropping systems and the characteristics of fields, and Radiant Earth Foundation for advice and guidance regarding machine learning best practices. We thank Manushi Trivedi, Sitian Xiong, and Tammy Woodard are thanked for their contributions to the underlying datasets and methods, and Michelle Gathigi, Omar Shehe, and Primoz Kovacic for support and management of the labelling efforts.
